---
title: "Problem Set 4"
author: "Neil Khurana"
date: "February 27, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(readr)
library(ggplot2)
library(lubridate)
library(readxl)
library(janitor)
library(gt)
```

## Question 1

```{r initialRead, include = FALSE}
# This will be the reference_csv for our polling data from NC 9th district
x <- read_csv("ps_4_elections-poll-nc09-3.csv")
```
[comment]: Specific filter for Dem voter and count rows which returns a number

There were `r x %>% filter(response == "Dem") %>% nrow()` respondents who supported the Democratic candidate.

[comment]: Filter of Rep and Undecided Voters, then subtract Und from Rep

There were `r x %>% filter(response == "Rep") %>% nrow() - x %>% filter(response == "Und") %>% nrow()` more respondents who favored the Republican candidate than who were Undecided.

[comment]: Filter out the unreadable values and keep only the values of gender and gender_combined that are differnt from each other. Then we count rows.

There are two gender variables (`gender` and `gender_combined`). There are `r x %>% filter(gender != "[DO NOT READ] Don't know/Refused", gender_combined != "[DO NOT READ] Don't know/Refused", gender != gender_combined) %>% nrow()` individuals for whom these variables have different values.

[comment]: This does a comparision of the races through filter. Count rows that differentiate

There are `r x %>% filter(race_eth == "White") %>% filter(race_eth != file_race_black) %>% nrow()` respondents listed as “White” under `race_eth` who are not listed as “White” under `file_race_black`.

```{r partyResponseTime, echo = FALSE}
#Finds the earliest time of response for a Republican Voter
min_r <- x %>% 
   filter(response == "Rep") %>%
    arrange(timestamp) %>% 
    slice(1) %>% 
  select(timestamp)
 
#Finds the earliest time of response for a Republican Voter
min_d <- x %>% 
  filter(response == "Dem") %>% 
  arrange(timestamp) %>% 
  slice(1) %>% 
  select(timestamp)
```

[comment]: Substracts the dem time from republican time and rounds to the nearest minute using round() function

The first `response` of Rep came `r round(as.numeric(min_r-min_d), digits = 0)` minutes (rounded to the nearest minute) before the first `response` of Dem.

## Question 2

```{r raceTable, echo = FALSE, message = FALSE}
race_chart <- x %>% 
  
  # The data is properly filtered to include only the race, response and weight values
  # Weight is used as our summing mechanism as explained in class to remove response bias from a poll
  filter(!is.na(response)) %>% 
  select(response, race_eth, final_weight) %>% 
  group_by(race_eth, response) %>% 
  summarize(total = sum(final_weight)) %>%   
  filter(race_eth != "[DO NOT READ] Don't know/Refused") %>% 
  
  # The data must be spread into a format where we can mutate the variables for calculation of percentage instead
  # 3rd party and all(overall) columns are no longer needed
  spread(key =  response, value = total, fill = 0) %>%  
  mutate(all = Dem + Rep + Und + `3`) %>% 
  mutate(Dem = Dem / all) %>% 
  mutate(Rep = Rep / all) %>% 
  mutate(Und = Und / all) %>% 
  select(-all, -`3`) %>%
  
  # Ungroup is very important since this is being transfered to a gt
  ungroup() %>% 
  
  # This mutate and arrange function orders race as we want it to be
  mutate(race_eth = factor(race_eth, levels = c("White", "Black", "Hispanic", "Asian", "Other"))) %>%
  arrange(race_eth) %>% 
  
  # All o's are NAs and this will be useful as seen in a future command
  na_if(0) %>%
  
  # Tibble is piped into gt()
  gt() %>% 
  
  # A title and source label is assigned
  tab_header(
    title = "Polling Results in North Carolina 9th Congressional District") %>% 
  tab_source_note(source_note = "Source: New York Times Upshot/Siena College 2018 Live Polls"
  ) %>% 
  
  # We want better labels than the variable names themselves. Race will now be a label and I prefer not to abbreviate
  cols_label(
    race_eth = "Race",
    Dem = "Democrat",
    Rep = "Republican",
    Und = "Undecided"
  ) %>%
  
  
  # The decimal values are converted to percents
  fmt_percent(columns = vars(Dem, Rep, Und),
              decimals = 0) %>% 
  
  # The NA gets replaced with a "-". This NA used to be a 0. 
  fmt_missing(columns = vars(Und), rows = 4) 

# Displays the table
race_chart


```

## Question 3

```{r educationPlot, echo = FALSE, message = FALSE}
educ_plot <- read_csv("ps_4_elections-poll-nc09-3.csv") %>%
  
  # Proper filters are put in place to get rid of NAs and meet the undergrad specifications
  select(educ, final_weight) %>% 
  filter(educ != "[DO NOT READ] Refused") %>%  
  

  # These are the levels of education we are interested. This from of mutate allows us to change the ordering we would like on our plot
  mutate(educ = factor(educ, levels = c("Grade school", "High school",
                                        "Some college or trade school", "Bachelors' degree",
                                        "Graduate or Professional Degree"))) %>%
  
  # Creation of a ggplot with geom_violin for the shape and geom_jitter for the scattered points. Opacity levels set
  ggplot(aes(x = educ, y = final_weight)) +
  geom_violin() +
  geom_jitter(alpha = 0.4, width = 0.2) +
  
  # Flips the display of axis for this graph
  coord_flip() +

  # Proper labeling of both axis along with subtitles and sources
  labs(title = "More Educated Matter Less in North Carolina 9th",
           subtitle = "Poll gives more weight to people who are less likely to participate in polls",
         x = NULL,
         y = "Weight Given to Respondent in Calculating Poll Results",
           caption = "New York Times Upshot/Siena College 2018 live polls") 

# Displays the plot
educ_plot
```

## Question 4

```{r genderBar, echo = FALSE, message = FALSE}

# In the following, we create a sideways bar chart to represent political party support by gender
# Each bar is separated by percentage in each gender. The appropriate weights are still kept in place
# This is a very interesting chart because we can determine if certain genders favor political parties

gender_table <- read_csv("ps_4_elections-poll-nc09-3.csv") %>%
  
  # Series of filter, select and summarize is performed so that we have columns of only gender party, and percentage
  filter(!is.na(response)) %>% 
  select(response, gender, final_weight) %>% 
  group_by(gender, response) %>% 
  summarize(total = sum(final_weight)) %>%   
  filter(gender != "[DO NOT READ] Don't know/Refused") %>% 
  
  # The table is spread so we can mutate new columns that break everything down by political party support
  # Removes exceptions like 3rd party
  spread(key =  response, value = total, fill = 0) %>%  
  mutate(all = Dem + Rep + Und + `3`) %>% 
  mutate(Dem = Dem / all * 100) %>% 
  mutate(Rep = Rep / all * 100) %>% 
  mutate(Und = Und / all * 100) %>% 
  select(-all, -`3`) %>%
  
# Data is regathered into form that can be plotted
  gather(party, pct, Dem:Und) 

# ggplot is initated with our previosuly created dataframe
gender_chart <- ggplot(gender_table, aes(x = party, y = pct, fill = party))

# Added properties to the ggplot, sideways bar and faceted by Gender
gender_chart + geom_col(position = "dodge2") +
    labs(x = NULL, y = "pct", fill = "party") +
    guides(fill = FALSE) + 
    coord_flip() + 
    facet_grid(~ gender) +
    
    # Each political party is assigned its respective color. 3rd party is green
    scale_fill_manual(values = c("blue", "red", "green")) +
    
    # All labels are included and title is centered
    ylab("Percentage of Voters") +
    xlab("Political Party") +
    labs(caption = "New York Times Upshot/Siena College 2018 live polls") +
    
    # Title is assigned and cenetered
    theme(plot.title = element_text(hjust = 0.5)) +
    ggtitle("Political Party Support Separated by Gender in NC 9th") +
    
    # An appropriate scale is selected to display the percentages
    scale_y_continuous(limits = c(0, 100)) 
```

# Colleagues
Jenny Li,
Mike Silva,
Tate Green,
Sabrina Chok,
Henry Zhu,
Albert Shin,
Roger Zhang


